---
title: "HW3, Onyeka Isamah"
output: github_document
---

```{r Setup}

library(tidyverse)
library(p8105.datasets)
```


# Problem 1:
The data set instacart has `r nrow(instacart) ` rows and `r ncol(instacart)` columns which include `order_id`, `product_id`, `add_to_cart_order`, `reordered user_id`, `eval_set`, `order_number`, `order_dow`, `order_hour_of_day`, `days_since_prior_order`, `product_name`, `aisle_id`, `department_id`, `aisle`, and `departmemt`. There are 134 aisles, the three top aisles that most products are order from include fresh fruits aisle (18726 bananas, ranked #1 in fresh fruit category), the packaged vegetable fruits aisle(Organic Baby Spiniach n = 9784 and ranked as #1 product within that aisle) and the milk aisle (organic whole milk n=4908).

```{r Instacart} 
data("instacart")
janitor::clean_names(instacart)
```

```{r Max Aisle}
max(instacart$aisle_id)
```

```{r top products by aisle}
top_products_by_aisle_df= instacart %>% select(
         everything()) %>% 
  group_by(aisle, department) %>% 
  count(
    product_name, name = "product_count") %>% 
  mutate(product_rank = min_rank(desc(product_count)))%>% 
 
   filter(product_rank <= 3)  %>% 
  arrange(product_rank) %>% arrange(desc(product_count)) 


top_products_by_aisle_df
```
 Plotting top products by aisle where product count > 10000
```{r top products plot}
top_products_by_aisle_df %>% 
filter(product_count >= 10000) %>% 
  ggplot(aes(x= product_name, y=product_count)) + geom_point(aes(color= aisle))
```


Table showing the three most popular items in each of the aisles “baking ingredients”, “dog food care”, and “packaged vegetables fruits”.

```{r 3 most popular}
pop3_df= (select(
  instacart, order_number, aisle,product_name) %>% 
  filter(aisle %in% c("baking ingredients", "dog food care", "packaged vegetables fruits")) %>% 
  
  group_by(aisle) %>% 
  count(
    product_name, name = "product_count") %>% 
    
  mutate(product_rank = min_rank(desc(product_count))) %>% 
    
   filter(product_rank <= 3)  %>% 
    
  arrange(product_rank) %>% arrange(desc(product_count)))

pop3_df
```

Make a table showing the mean hour of the day at which Pink Lady Apples and Coffee Ice Cream are ordered on each day of the week; format this table for human readers (i.e. produce a 2 x 7 table).

```{r mean hod}
mean_hod = instacart %>% drop_na() %>% 
  select(product_name,order_hour_of_day, order_dow)  %>% 
 mutate(
    
    order_dow = 
      recode_factor(order_dow,
        "0" = "Sunday", "1" = "Monday", "2" = "Tuesday", "3" = "Wednesday", "4"= "Thursday", "5" = "Friday", "6" = "Saturday")) %>% 
  group_by(product_name, order_dow) %>% 
  summarise(meanhour = mean(order_hour_of_day)) %>% 
filter(product_name %in% c( "Pink Lady Apples", "Coffee Ice Cream")) %>% 
  pivot_wider(names_from = order_dow,  values_from = meanhour) 

mean_hod
```

#Problem 2

```{r import}
library(p8105.datasets)
data("brfss_smart2010")

brfss_smart2010

```

```{r cleaning}

newbrfss_df <- brfss_smart2010 %>% separate(Locationdesc,into = c ("state", "County")) %>% 
select(Response, Topic, everything()) %>% 
  mutate(Response = factor(Response, levels = c("Poor", "Fair", "Good", "Very good", "Excellent"))) %>% 
  filter(Topic == "Overall Health") %>% janitor::clean_names() 
newbrfss_df
```


#In 2002, which states were observed at 7 or more locations? 
5 states 

```{r 2002}
brfss2002 = newbrfss_df %>% 
  filter(year =="2002") %>% 
  group_by(state) %>% 
  distinct(county) %>% 
  count(state) %>% 
  filter(n>=7) %>% 
  select(state)
brfss2002
```

14 states in 2010
```{r 2010}
brfss2010 = newbrfss_df %>% 
  filter(year =="2010") %>% 
  group_by(state) %>% 
  distinct(county) %>% 
  count(state) %>% 
  filter(n>=7) %>% 
  select(state)
brfss2010
```


Construct a dataset that is limited to Excellent responses, and contains, year, state, and a variable that averages the data_value across locations within a state. Make a “spaghetti” plot of this average value over time within a state (that is, make a plot showing a line for each state across years

```{r Excellent dataset and plot}
newbrfss_df %>% 
  select(year, locationabbr, response, data_value, geo_location) %>% 
  filter(response == "Excellent") %>% 
  group_by(locationabbr,year) %>% 
  mutate(avg_data = mean(data_value)) %>% 
  select(year, locationabbr, avg_data) %>% 
  
  
  ggplot(aes(x=year, y = avg_data, color= locationabbr)) + 
  geom_line() + theme(legend.position = "right") 
```


Make a two-panel plot showing, for the years 2006, and 2010, distribution of data_value for responses (“Poor” to “Excellent”) among locations in NY State.
```{r facet NY}
facetbrfss <- brfss_smart2010 %>% 
select(Response, Topic, everything()) %>% 
  mutate(Response = factor(Response, levels = c("Poor", "Fair", "Good", "Very good", "Excellent"))) %>% 
  filter(Topic == "Overall Health") %>% janitor::clean_names() 
newbrfss_df

facetbrfss%>% 
  filter(locationabbr == "NY", year %in% c("2006", "2010")) %>% 
  group_by(year,locationdesc) %>% 
  ggplot(aes(x=response, y = data_value, group = locationdesc, color =locationdesc))+ geom_line()+facet_grid(.~year)
```


